A Tale of Two Dictionary Learning Problems

### 2020-08-05 manuscripts/thesis/proofs.tex
chapters
new intro

\chapter{Definitions and Proofs}\label{DUT} % of Theorems~\ref{DeterministicUniquenessTheorem} and ~\ref{SLCopt}}\label{DUT}

\section{Introduction}

@@

\section{The main lemma}

An extension of the proof to the general case $k < m$ requires some additional tools to derive the general expression \eqref{Cdef1} for $C_1$. These include a generalized notion of distance (Def.~\ref{dDef}) and angle (Def.~\ref{FriedrichsDefinition}) between subspaces as well as a stability result in combinatorial matrix analysis (Lem.~\ref{MainLemma}), which is the main lemma doing most of the heavy lifting in the proof of Thm.~\ref{DeterministicUniquenessTheorem}.

@@

The required stability result in combinatorial matrix analysis is the following.

@@

Several results in the literature had to be combined to extend the guarantees derived in \cite{Hillar15} into the noisy regime. The main challenge was to generalize Lem.~\ref{MainLemma} to the case where the $k$-dimensional subspaces spanned by corresponding submatrices $\mathbf{A}_S$ and $\mathbf{B}_{\pi(S)}$ are assumed to be``close" but not identical. Unlike in \cite{Hillar15} where an inductive argument could be applied to the noiseless case, here it had to be explicitly demonstrated that this proximity relation is propagated through iterated intersections right down to the spans of the dictionary elements themselves. Lem.~\ref{DistanceToIntersectionLemma} was designed to encapsulate this fact, proven by appeal to a convergence guarantee for an alternating projections algorithm first proposed by von Neumann. This result, combined with a little known fact \eqref{eqdim} about the distance metric between subspaces, make up the more obscure components of the deduction.

@@

As it seemed to benefit a reviewer of this work, a brief discussion on how the deterministic sample complexity $N = |\mathcal{H}\left[|(k-1){m \choose k} + 1\right]$ derived here compares to those listed in the the top two rows listed in Table I of \cite{Hillar15} may be in order. To be clear, the theory developed here is strictly more general, since $\mathcal{H}$ can always be taken to be ${[m] \choose k}$. The point of difference in this comparison is the assumed set of supports for sparse codes, which is always ${[m] \choose k}$ in \cite{Hillar15}, whereas here it can be assumed to be any regular $k$-uniform hypergraph that satisfies the SIP. By row, %, assuming for our results that sufficient data have been sampled from all supports in a hypergraph H satisfying the SIP:
\begin{enumerate}
\item[I.] The result here improves upon the listed $k{m \choose k}^2$ by an exponential factor, since for every $k < m$ there exists a regular $k$-uniform hypergraph $\mathcal{H}$ with $|\mathcal{H}| = m$ satisfying the SIP. 
\item[II.] The authors in \cite{Hillar15} have applied measure-theoretic arguments to achieve $(k+1){m \choose k}$ with almost-certainty (i.e. with probability one), a factor of $m$ reduction over that for which certainty can alternatively be guaranteed here.    
%\item[III.] A direct comparison cannot be made because it is an open question as to the probability that a random set of supports satisfies the SIP. This would be an interesting problem for the community to solve; regardless, the result of \cite{Hillar15} is still implied by the more general theory given here.
%\item[IV.] The result here is $|\mathcal{H}\left[|(k-1){m \choose k} + 1\right]$ with probability one compared to $(k+1){m \choose k}$ with probability one. In this case the sample complexity stated in \cite{Hillar15} is marginally better, but it relies on a noise-free technicality unique to their assumption $\mathcal{H} = {m \choose k}$. %differs in flavor: their $\mathbf{x_i}$ are assumed to be distributed as $(k+1)$ samples per support for every support of size $k \in [m]$, whereas here it is assumed that $(k-1){m \choose k}$ codes $\mathbf{x_i}$ are distributed over each support in some hypergraph $\mathcal{H}$ satisfying the SIP. 
%\item[V.] Again, a direct comparison cannot be made for the reason stated for row III.
\end{enumerate}

>>>>

\chapter{Proofs}\label{DUT} % of Theorems~\ref{DeterministicUniquenessTheorem} and ~\ref{SLCopt}}\label{DUT}

@@

\section{Stating the main lemma}

An extension of the proof to the general case $k < m$ requires some additional tools to derive the general expression \eqref{Cdef1} for $C_1$. These include a generalized notion of distance (Def.~\ref{dDef}) and angle (Def.~\ref{FriedrichsDefinition}) between subspaces as well as a stability result in combinatorial matrix analysis (Lem.~\ref{MainLemma}), which contains most of the complexity of the proof of Thm.~\ref{DeterministicUniquenessTheorem}.

@@

The required stability result in combinatorial matrix analysis is the following. For expository purposes, the proof of this fact is relegated to the Appendix.

@@

Several results in the literature had to be combined to extend the guarantees derived in \cite{Hillar15} into the noisy regime. The main challenge was to generalize Lem.~\ref{MainLemma} to the case where the $k$-dimensional subspaces spanned by corresponding submatrices $\mathbf{A}_S$ and $\mathbf{B}_{\pi(S)}$ are assumed to be``close" but not identical. Referring now to the proof in the Appendix, the situation is unlike that in \cite{Hillar15}, where an inductive argument could be applied to the noiseless case. Rather, here it has to be explicitly demonstrated that this proximity relation is propagated through iterated intersections right down to the spans of the dictionary elements themselves. Lem.~\ref{DistanceToIntersectionLemma} was designed to encapsulate this fact, proven by appeal to a convergence guarantee for an alternating projections algorithm first proposed by von Neumann. This result, combined with a little known fact \eqref{eqdim} about the distance metric between subspaces, make up the more obscure components of the deduction.

@@

\section{Appendix: Proving the main lemma}\label{proofs}

We shall prove Lem.~\ref{MainLemma} after the following auxiliary lemmas.  % stating and proving some auxiliary lemmas. 
%First, note that since $\|\mathbf{A}\mathbf{x}\|_2 \leq \max_j\|\mathbf{A}_j\|_2\|\mathbf{x}\|_1$ and $\|\mathbf{x}\|_1 \leq \sqrt{k} \|\mathbf{x}\|_2$ for $k$-sparse $\mathbf{x}$, by \eqref{Ldef} we have the following frequently applied inequality:
%\begin{align}\label{delrho}
%L_{k'}(\mathbf{A}) \leq  \max_j\|\mathbf{A}_j\|_2 \ \ \text{for all $k' \leq 2k$}.
%\end{align}

% and then sketch the proofs of Thm.~\ref{robustPolythm} and Cor.~\ref{ProbabilisticCor}.

%\begin{lemma}\label{spanIntersectionLemma}
%Let $\mathbf{M} \in \mathbb{R}^{n \times m}$. If every $2k$ columns of $\mathbf{M}$ are linearly independent, then for any $F \subseteq {[m] \choose k}$:
%\begin{align*}
%\text{\rmfamily span}\{\mathbf{M}_{\cap F}\}  = \bigcap_{S \in F} \text{\rmfamily span}\{\mathbf{M}_S\}.
%\end{align*}
%\end{lemma}
%\begin{proof}
%Using induction on $|F|$, it is enough to prove the lemma when $|F| = 2$; but this case follows directly from the assumption.
%\end{proof}

\begin{lemma}\label{spanIntersectionLemma}
If $f: V \to W$ is injective, then $f\left(\cap_{i=1}^\ell V_i \right) =  \cap_{i=1}^\ell f\left(V_i\right)$ for any $V_1, \ldots, V_\ell \subseteq V$. ($f(\emptyset):=\emptyset$.)
%In particular, if for some $E \in 2^{[m]}$ the map $M \in \mathbb{R}^{n \times m}$ is injective on $\cup_{S \in \mathcal{H}} \text{span}\{e_i\}_{i \in S}$ then $\text{span}\{ M_{\cap_{S \in \mathcal{H}} S} \} = \cap_{S \in \mathcal{H}} \text{span}\{M_S\}$.
\end{lemma}
\begin{proof}
By induction, it is enough to prove the case $\ell = 2$. Clearly, for any map $f$, if $w \in f(U \cap V)$ then $w \in f(U)$ and $w \in f(V)$; hence, $w \in f(U) \cap f(V)$. If $w \in f(U) \cap f(V)$, then $w \in f(U)$ and $w \in f(V)$; thus, $w = f(u) = f(v)$ for some $u \in U$ and $v \in V$, implying $u = v$ by injectivity of $f$. It follows that $u \in U \cap V$ and $w \in f(U \cap V)$.
\end{proof}
In particular, if a matrix $\mathbf{A}$ satisfies $L_{2\mathcal{H}}(\mathbf{A}) > 0$, then taking $V$ to be the union of subspaces consisting of vectors with supports in $2\mathcal{H}$, we have $\bm{\mathcal{A}}_{\cap \mathcal{G}} = \cap \bm{\mathcal{A}}_\mathcal{G}$ for all $\mathcal{G} \subseteq \mathcal{H}$.
% \vspace{-.4 cm}

\begin{lemma}\label{DistanceToIntersectionLemma}
Let $\mathcal{V} = \{V_i\}_{i=1}^k$ be a set of two or more subspaces of $\mathbb{R}^m$, and set $V = \cap \mathcal{V} $. For  $\mathbf{u} \in \mathbb{R}^m$, we have (recall Defs.~\ref{dDef}~\&~\ref{FriedrichsDefinition}):
\begin{align}\label{DTILeq}
\text{\rmfamily dist}(\mathbf{u}, V) \leq \frac{1}{1 - \xi(\mathcal{V})} \sum_{i=1}^k \text{\rmfamily dist}(\mathbf{u}, V_i).
\end{align}
\end{lemma}
\begin{proof} 
% When $V = \{\mathbf{0}\}$, the result is trivial, so suppose otherwise.  
Recall the projection onto the subspace $V \subseteq \mathbb{R}^m$ is the mapping $\Pi_V: \mathbb{R}^m \to V$ that associates with each $\mathbf{u}$ its unique nearest point in $V$; i.e., $\|\mathbf{u} - \Pi_V\mathbf{u}\|_2 = \text{\rmfamily dist}(\mathbf{u}, V)$.
By repeatedly applying the triangle inequality, we have:
\begin{align}\label{f}
\|\mathbf{u} - &\Pi_V\mathbf{u}\|_2 
\leq \|\mathbf{u} - \Pi_{V_k} \mathbf{u}\|_2 + \|\Pi_{V_k}  \mathbf{u} - \Pi_{V_k}\Pi_{V_{k-1}}\mathbf{u}\|_2 \nonumber \\
&\ \ \ \ \ \ \ \ \ \ \ + \cdots + \|\Pi_{V_k} \Pi_{V_{k-1}}\cdots \Pi_{V_1} \mathbf{u} - \Pi_V \mathbf{u}\|_2 \nonumber \\
&\leq  \sum_{\ell=1}^k \|\mathbf{u} - \Pi_{V_{\ell}} \mathbf{u}\|_2 
+ \|(\Pi_{V_k}\cdots\Pi_{V_{1}} - \Pi_V) \mathbf{u}\|_2,
\end{align}
where we have also used that the spectral norm of the orthogonal projections $\Pi_{V_{\ell}}$ satisfies $\|\Pi_{V_{\ell}}\|_2 \leq 1$ for all $\ell$.

It remains to bound the second term in \eqref{f} by $\xi(\mathcal{V}) \|\mathbf{u} - \Pi_V\mathbf{u}\|_2$. First, note that $\Pi_{V_\ell} \Pi_V = \Pi_V$ and $\Pi_V^2 = \Pi_V$, so we have $\|(\Pi_{V_k} \cdots \Pi_{V_1} - \Pi_V) \mathbf{u} \|_2 
= \| ( \Pi_{V_k} \cdots\Pi_{V_1} - \Pi_V ) (\mathbf{u} - \Pi_V\mathbf{u})\|_2$. % \leq z\|\mathbf{u} - \Pi_V\mathbf{u}\|_2$.
Consequently, inequality \eqref{DTILeq} follows from \cite[Thm.~9.33]{Deutsch12}:
\begin{align}
\|\Pi_{V_k}\Pi_{V_{k-1}}\cdots\Pi_{V_1} \mathbf{x} - \Pi_V\mathbf{x}\|_2 \leq z \|\mathbf{x}\|_2, \ \ \text{for all } \mathbf{x},
\end{align}
with \mbox{$z^2= 1 - \prod_{\ell =1}^{k-1}(1-z_{\ell}^2)$} and \mbox{$z_{\ell} = \cos\theta\left(V_{\ell}, \cap_{s=\ell+1}^k V_s\right)$} (recall $\theta$ from Def.~\ref{FriedrichsDefinition}), after substituting $\xi(\mathcal{V})$ for $z$ and rearranging terms.
\end{proof}
\begin{lemma}\label{NonEmptyLemma} 
Fix an $r$-regular hypergraph $\mathcal{H} \subseteq 2^{[m]}$ satisfying the SIP. If the map $\pi: \mathcal{H} \to 2^{[\overline m]}$ has $\sum_{S \in \mathcal{H}} |\pi(S)| \geq \sum_{S \in \mathcal{H}} |S|$ and:
\begin{align}\label{cond}
	|\cap \pi(\mathcal{G})| \leq |\cap \mathcal{G} |,\ \ \   \text{for } \mathcal{G} \in {\mathcal{H} \choose r} \cup {\mathcal{H} \choose r+1},
\end{align}
%
then $\overline m \geq m$; and if $\overline m  (r-1) < mr$, the map $i \mapsto \cap_{S \in \sigma(i)} \pi(S)$ is an injective function to $[\overline m]$ from some $J \subseteq [m]$ of size $m - (r-1)(\overline m - m)$ (recall $\sigma$ from Def.~\ref{sip}).  %In particular, if $\overline m = m$ then $\pi$ is induced by a permutation on $[m]$.
\end{lemma}

\begin{proof}
Consider the following set: $T_1 := \{(i, S): i \in \pi(S), S \in \mathcal{H}\}$, which numbers $|T_1| = \sum_{S \in \mathcal{H}} |\pi(S)| \geq \sum_{S \in \mathcal{H}} |S| = \sum_{i \in [m]} \deg_\mathcal{H}(i) = mr$ by $r$-regularity of $\mathcal{H}$. Note that $|T_1| \leq \overline m r$; otherwise, pigeonholing the tuples of $T_1$ with respect to their $\overline m$ possible first elements would imply that more than $r$ of the tuples in $T_1$ share the same first element. This cannot be the case, however, since then some $\mathcal{G} \in {\mathcal{H} \choose r+1}$ formed from any $r+1$ of their second elements would satisfy $\cap \pi(\mathcal{G}) \neq 0$; hence, $|\cap \mathcal{G}| \neq 0$ by \eqref{cond}, contradicting $r$-regularity of $\mathcal{H}$. It follows that $\overline m \geq m$.

Suppose now that $\overline m (r-1) < mr$, so that $p := mr - \overline m (r-1)$ is positive and $|T_1| \geq \overline m (r - 1) + p$. Pigeonholing $T_1$ into $[\overline m]$ again, there are at least $r$ tuples in $T_1$ sharing some first element; that is, for some $\mathcal{G}_1 \subseteq \mathcal{H}$ of size $|\mathcal{G}_1| \geq r$, we have $|\cap \pi(\mathcal{G}_1)| \geq 1$ and (by \eqref{cond}) $|\cap \mathcal{G}_1| \geq 1$. Since no more than $r$ tuples of $T_1$ can share the same first element, we in fact have $|\mathcal{G}_1| = r$. It follows by $r$-regularity that $\mathcal{G}_1$ is a star of $\mathcal{H}$; hence, $|\cap \mathcal{G}_1| = 1$ by the SIP and $|\cap \pi(\mathcal{G}_1)|  = 1$ by \eqref{cond}.

If $p=1$, then we are done. Otherwise, define $T_2 := T_1 \setminus \{(i,S) \in T_1: i = \cap \pi(\mathcal{G}_1)\}$, which contains $|T_2| = |T_1| - r \geq (\overline m - 1)(r-1) + (p-1)$ ordered pairs having $\overline m - 1$ distinct first indices. Pigeonholing $T_2$ into $[\overline m - 1]$ and repeating the above arguments produces the star $\mathcal{G}_2 \in {\mathcal{H} \choose r}$ with intersection $\cap \mathcal{G}_2$ necessarily distinct (by $r$-regularity) from $\cap \mathcal{G}_1$. Iterating this procedure $p$ times in total yields the stars $\mathcal{G}_i$ for which $\cap\mathcal{G}_i \mapsto \cap \pi(\mathcal{G}_i)$ defines an injective map to $[\overline m]$ from $J = \{\cap \mathcal{G}_1, \ldots, \cap \mathcal{G}_p\} \subseteq [m]$.
\end{proof}

\begin{proof}[Proof of Lem.~\ref{MainLemma}]
%We will show that the bound \eqref{GapUpperBound} trickles through the intersection semi-lattices of $\{\bm{\mathcal{A}}_S\}_{S \in \mathcal{H}}$ and $\{\bm{\mathcal{B}}_{\overline S}\}_{\overline S \in \pi(\mathcal{H})}$ to yield \eqref{MainLemmaBPD} by virtue of the SIP.  
Let us begin by showing that $\dim(\bm{\mathcal{B}}_{\pi(S)}) = \dim(\bm{\mathcal{A}}_S)$ for all $S \in \mathcal{H}$. Note that since $\|\mathbf{A}\mathbf{x}\|_2 \leq \max_j\|\mathbf{A}_j\|_2\|\mathbf{x}\|_1$ and $\|\mathbf{x}\|_1 \leq \sqrt{k} \|\mathbf{x}\|_2$ for all $k$-sparse $\mathbf{x}$, by \eqref{Ldef} we have $L_2(\mathbf{A}) \leq \max_j\|\mathbf{A}_j\|_2$ and therefore (as $0 \leq \xi < 1$) the right-hand side of \eqref{GapUpperBound} is less than one. From \eqref{dimLem}, we have $|\pi(S)| \geq \dim(\bm{\mathcal{B}}_{\pi(S)}) \geq \dim(\bm{\mathcal{A}}_S) = |S|$, the final equality holding by injectivity of $\mathbf{A}_S$. As $|\pi(S)| = |S|$, the claim follows. Note, therefore, that $\mathbf{B}_{\pi(S)}$ has full-column rank for all $S \in \mathcal{H}$.

We shall next demonstrate that \eqref{cond} holds. Fixing $\mathcal{G} \in {\mathcal{H} \choose r} \cup {\mathcal{H} \choose r+1}$, it suffices to show that $d(\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})}, \bm{\mathcal{A}}_{\cap \mathcal{G}} ) < 1$, since by \eqref{dimLem} we then have $|\cap \pi(\mathcal{G})| = \dim(\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})}) \leq \dim(\bm{\mathcal{A}}_{\cap \mathcal{G}}) = |\cap \mathcal{G}|$, with equalities from the full column-ranks of $\mathbf{A}_{S}$ and $\mathbf{B}_{\pi(S)}$ for all $S \in \mathcal{H}$.\footnote{Note that if ever $\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})} \neq /bfseries 0$ while $\cap \mathcal{G} = \emptyset$, we would have $d(\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})}, \bm 0 ) = 1$. However, that leads to a contradiction.} Observe that $d(\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})}, \bm{\mathcal{A}}_{\cap \mathcal{G}}  ) 
\leq d\left( \cap \bm{\mathcal{B}}_{\pi(\mathcal{G})}, \cap \bm{\mathcal{A}}_\mathcal{G} \right)$ by \eqref{UsubU}, since trivially $\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})} \subseteq \cap \bm{\mathcal{B}}_{\pi(\mathcal{G})}$ and also $\bm{\mathcal{A}}_{\cap \mathcal{G}} = \cap \bm{\mathcal{A}}_\mathcal{G}$ by Lem.~\ref{spanIntersectionLemma}. Recalling Def.~\ref{dDef} and applying Lem.~\ref{DistanceToIntersectionLemma} yields:
\begin{align}
d\left( \cap \bm{\mathcal{B}}_{\pi(\mathcal{G})}, \cap \bm{\mathcal{A}}_\mathcal{G} \right)
&\leq \max_{\mathbf{u} \in \cap \bm{\mathcal{B}}_{\pi(\mathcal{G})}, \ \|\mathbf{u}\|_2 \leq 1} \sum_{S \in \mathcal{G}} \frac{ \text{\rmfamily dist}\left( \mathbf{u},\bm{\mathcal{A}}_{S} \right) }{ 1 - \xi( \bm{\mathcal{A}}_\mathcal{G} ) } \nonumber \\
&= \sum_{S \in \mathcal{G}} \frac{ d\left( \cap \bm{\mathcal{B}}_{\pi(\mathcal{G})},\bm{\mathcal{A}}_{S} \right) }{ 1 - \xi( \bm{\mathcal{A}}_\mathcal{G} ) }, \nonumber
\end{align}
passing the maximum through the sum.
Since $\cap \bm{\mathcal{B}}_{\pi(\mathcal{G})} \subseteq \bm{\mathcal{B}}_{\pi(S)}$ for all $S \in \mathcal{G}$, by \eqref{UsubU} the numerator of each term in the sum above is bounded by \mbox{$d\left( \bm{\mathcal{B}}_{\pi(S)},\bm{\mathcal{A}}_{S} \right) = d\left(\bm{\mathcal{A}}_{S}, \bm{\mathcal{B}}_{\pi(S)} \right) \leq \varepsilon$}, with the equality from \eqref{eqdim} since $\dim(\bm{\mathcal{B}}_{\pi(S)}) = \dim(\bm{\mathcal{A}}_S)$. Thus, altogether:
\begin{align}\label{last}
d(\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})}, \bm{\mathcal{A}}_{\cap \mathcal{G}} )
\leq \frac{|\mathcal{G}| \varepsilon}{1 - \xi( \bm{\mathcal{A}}_\mathcal{G} )}
\leq \frac{C_2 \varepsilon}{\max_j\|\mathbf{A}_j\|_2},
\end{align}
recalling the definition of $C_2$ in \eqref{Cdef2}. Lastly, since $C_2 \varepsilon < L_2(\mathbf{A}) \leq \max_j\|\mathbf{A}_j\|_2$, we have $d(\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})}, \bm{\mathcal{A}}_{\cap \mathcal{G}} ) \leq 1$ and therefore \eqref{cond} holds.

%Keeping in mind that $d(U',V) \leq d(U,V)$ whenever $U' \subseteq U$ and applying (in order) Lem.~\ref{spanIntersectionLemma}, Lem.~\ref{DistanceToIntersectionLemma}, \eqref{eqdim}, \eqref{GapUpperBound}, and \eqref{Cdef2} gives:
%\begin{align}\label{randoml}
%d(&\bm{\mathcal{B}}_{\cap \pi(\mathcal{G})}, \bm{\mathcal{A}}_{\cap \mathcal{G}}  ) 
%\leq d\left( \cap \bm{\mathcal{B}}_{\pi(\mathcal{G})}, \cap \bm{\mathcal{A}}_\mathcal{G} \right)
%\leq \sum_{T \in \mathcal{G}} \frac{ d\left( \cap \bm{\mathcal{B}}_{\pi(\mathcal{G})},\bm{\mathcal{A}}_{T} \right) }{ 1 - \xi( \bm{\mathcal{A}}_\mathcal{G} ) } \nonumber \\
%&\leq \sum_{T \in \mathcal{G}} \frac{ d\left( \bm{\mathcal{B}}_{\pi(T)},\bm{\mathbf{A}}_{T} \right) }{ 1 - \xi( \bm{\mathcal{A}}_\mathcal{G} ) }
%\leq \frac{|\mathcal{G}| \varepsilon}{1 - \xi( \bm{\mathcal{A}}_\mathcal{G} )} 
%\leq \frac{C_2 \varepsilon}{\max_j\|\mathbf{A}_j\|_2},
%\end{align}
%
%which is less than one, since $C_2 \varepsilon < L_2(\mathbf{A}) \leq \max_j\|\mathbf{A}_j\|_2$.

Applying Lem.~\ref{NonEmptyLemma}, the association $i \mapsto \cap_{S \in \sigma(i)} \pi(S)$ is an injective map $\overline \pi: J \to [\overline m]$ for some $J \subseteq [m]$ of size $m - (r-1)(\overline m - m)$, and $\mathbf{B}_{\overline \pi(i)} \neq \mathbf{0}$ for all $i \in J$ since the columns of $\mathbf{B}_{\pi(S)}$ are linearly independent for all $S \in \mathcal{H}$. Letting $\overline \varepsilon := C_2 \varepsilon / \max_i \|\mathbf{A}_i\|_2$, it follows from \eqref{eqdim} and \eqref{last} that $d\left( \bm{\mathcal{A}}_i, \bm{\mathcal{B}}_{\overline \pi(i)} \right) = d\left(\bm{\mathcal{B}}_{\overline \pi(i)},  \bm{\mathcal{A}}_i \right)  \leq \overline \varepsilon$ for all $i \in J$. %Fixing $\overline \varepsilon = C_2\varepsilon$ and letting $c_i = \|\mathbf{A}_i\|_2^{-1}$, we thus have that for each $\mathbf{e}_i \in \mathbb{R}^m$ with $i \in J$ there exists some $\overline{c}_i \in \mathbb{R}$ such that $\|c_i\mathbf{A}\mathbf{e}_i - \overline{c}_i \mathbf{B}\mathbf{e}_{\overline \pi(i)}\|_2 \leq \overline \varepsilon < L_2(\mathbf{A}) \min_{i\in J} |c_i|$. But this is exactly the supposition in \eqref{1D}, and the result follows from the case $k=1$ in Sec.~\ref{DUT} applied to the submatrix $\mathbf{A}_J$. 
Setting $c_i := \|\mathbf{A}_i\|_2^{-1}$ so that $\|c_i\mathbf{Ae}_i\|_2 = 1$, by Def.~\ref{dDef} for all $i \in J$:
\begin{align*}
\min_{\overline c_i \in \mathbb{R}} \|c_i\mathbf{Ae}_i - \overline c_i \mathbf{Be}_{\overline \pi(i)} \|_2
\leq d\left( \bm{\mathcal{A}}_i, \bm{\mathcal{B}}_{\overline \pi(i)} \right)
\leq \overline \varepsilon,
%&= \max_{\bm{u} \in \bm{\mathcal{A}}_i, \|\bm{u}\|_2 \leq 1} \text{dist}\left(\bm{u}, \bm{\mathcal{B}}_{\overline \pi(i)} \right) \\
%&\geq  \text{dist}\left(c_i\bm{Ae}_i, \mathbf{Be}_{\overline \pi(i)} \right) \\
%&= \min_{\overline c_i \in \mathbb{R}} \|c_i\bm{Ae}_i - \overline c_i \mathbf{Be}_{\overline \pi(i)} \|_2
\end{align*}
%
for $\overline \varepsilon < L_2(\mathbf{A})\min_{i \in [m]}|c_i|$. But this is exactly the supposition in \eqref{1D}, with $J$ and $\overline \varepsilon$ in place of $[m]$ and $\varepsilon$, respectively. The same arguments of the case $k=1$ in Sec.~\ref{DUT} can then be made to show that for any $\overline m \times \overline m$ permutation and invertible diagonal matrices $\mathbf{P}$ and $\mathbf{D}$ with, respectively, columns $\mathbf{e}_{\pi(i)}$ and $\frac{\overline{c}_i}{c_i}\mathbf{e}_i$ for $i \in J$ (otherwise, $\mathbf{e}_{i}$ for $i \in [\overline{m}] \setminus J$), we have $\|\mathbf{A}_i - (\mathbf{B}\mathbf{PD})_i \|_2 \leq \overline  \varepsilon / |c_i|  \leq C_2 \varepsilon$ for all $i \in J$.
\end{proof}

### 2020-08-05 manuscripts/thesis/results.tex
new intro

\section{Introduction}

TODO

@@

\section{Combinatorial designs}

@@

\section{Uniqueness Theorems}

@@

Let us conclude this chapter with some comments on the optimality of these results.  The linear scaling for $\varepsilon$ in \eqref{epsdel} is essentially optimal (e.g., see \cite{arias2013fundamental}), but a basic open problem remains: how many samples are necessary to determine the sparse coding model? 
These results demonstrate that sparse codes $\mathbf{x}_i$ drawn from only a polynomial number of $k$-dimensional subspaces permit stable identification of the generating dictionary $\mathbf{A}$. 
This lends some legitimacy to the use of the model in practice, where data in general are unlikely (if ever) to exhibit the exponentially many possible $k$-wise combinations of dictionary elements required by (to my knowledge) all previously published results.

Consequently, if $k$ is held fixed or if the size of the support set of reconstructing codes is polynomial in $\overline m$ and $k$, then a practical (polynomial) amount of data suffices to identify the dictionary.\footnote{In the latter case, a reexamination of the pigeonholing argument in the proof of Thm.~\ref{DeterministicUniquenessTheorem} requires a polynomial number of samples distributed over a polynomial number of supports.} Reasons to be skeptical that this holds in general, however, can be found in \cite{tillmann2014computational, tillmann2015computational}. Even so, in the next chapter it will be discussed how probabilistic guarantees can in fact be made for any number of available samples.

>>>>

@@

\section{Definitions}

@@

\section{Uniqueness theorems}

@@

It is befitting to conclude this chapter with some comments on the optimality of these results.  The linear scaling for $\varepsilon$ in \eqref{epsdel} is essentially optimal (e.g., see \cite{arias2013fundamental}), but a basic open problem remains: how many samples are necessary to determine the sparse coding model? 
These results demonstrate that sparse codes $\mathbf{x}_i$ drawn from only a polynomial number of $k$-dimensional subspaces permit stable identification of the generating dictionary $\mathbf{A}$. 
This lends some legitimacy to the use of the model in practice, where data in general are unlikely (if ever) to exhibit the exponentially many possible $k$-wise combinations of dictionary elements required by (to my knowledge) all previously published results.

Consequently, if $k$ is held fixed or if the size of the support set of reconstructing codes is polynomial in $\overline m$ and $k$, then a practical (polynomial) amount of data suffices to identify the dictionary.\footnote{In the latter case, a reexamination of the pigeonholing argument in the proof of Thm.~\ref{DeterministicUniquenessTheorem} requires a polynomial number of samples distributed over a polynomial number of supports.} Reasons to be skeptical that this holds in general, however, can be found in \cite{tillmann2014computational, tillmann2015computational}. Even so, the next chapter contains a discussion on how probabilistic guarantees can in fact be made for any number of available samples.

As it seemed to benefit a reviewer of this work, some clarification may be in order on how the deterministic sample complexity $N = |\mathcal{H}\left[|(k-1){m \choose k} + 1\right]$ given here compares to those listed in the the top two rows listed in Table I of \cite{Hillar15}. To be clear, the theory developed here (Cor.~\ref{DeterministicUniquenessCorollary}) is strictly more general, since $\mathcal{H}$ can always be taken to be ${[m] \choose k}$. The point of difference in this comparison is the assumed set of supports for sparse codes, which is always ${[m] \choose k}$ in \cite{Hillar15}, whereas here it can be assumed to be any regular $k$-uniform hypergraph that satisfies the SIP. By row, %, assuming for our results that sufficient data have been sampled from all supports in a hypergraph H satisfying the SIP:
\begin{enumerate}
\item[I.] The result here improves upon the listed $k{m \choose k}^2$ by an exponential factor, since for every $k < m$ there exists a regular $k$-uniform hypergraph $\mathcal{H}$ with $|\mathcal{H}| = m$ satisfying the SIP. 
\item[II.] The authors in \cite{Hillar15} have applied measure-theoretic arguments to achieve $(k+1){m \choose k}$ with almost-certainty (i.e. with probability one), a factor of $m$ reduction over that for which certainty can alternatively be guaranteed here.    
%\item[III.] A direct comparison cannot be made because it is an open question as to the probability that a random set of supports satisfies the SIP. This would be an interesting problem for the community to solve; regardless, the result of \cite{Hillar15} is still implied by the more general theory given here.
%\item[IV.] The result here is $|\mathcal{H}\left[|(k-1){m \choose k} + 1\right]$ with probability one compared to $(k+1){m \choose k}$ with probability one. In this case the sample complexity stated in \cite{Hillar15} is marginally better, but it relies on a noise-free technicality unique to their assumption $\mathcal{H} = {m \choose k}$. %differs in flavor: their $\mathbf{x_i}$ are assumed to be distributed as $(k+1)$ samples per support for every support of size $k \in [m]$, whereas here it is assumed that $(k-1){m \choose k}$ codes $\mathbf{x_i}$ are distributed over each support in some hypergraph $\mathcal{H}$ satisfying the SIP. 
%\item[V.] Again, a direct comparison cannot be made for the reason stated for row III.
\end{enumerate}
